Tesseroids: forward modeling gravitational fields in spherical coordinates

### 2015-03-07 manuscript.tex
Replaced textit with emph

\pdfoutput=1
\documentclass[paper,twocolumn]{geophysics}
%\documentclass[manuscript]{geophysics}
%\documentclass[manuscript,endfloat]{geophysics}
%\documentclass[twocolumn]{article}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\newcommand{\mitbf}[1]{\mathbf{{#1}}}

@@

We have implemented a modified version of the adaptive discretion of
\citet{Li2011} into the open-source software package \textit{Tesseroids}.
The software uses the Cartesian formula of \citet{Grombein2013} for improved
performance and robustness.
Previous versions of \textit{Tesseroids} have been used by, e.g.,
\citet{Alvarez2012, Bouman2013, Bouman2013a, Mariani2013, Braitenberg2014,
Braitenberg2011, Fullea2014}.
This article describes the software \textit{Tesseroids} and the implementation
of our modified adaptive discretization algorithm.
We also present a numerical investigation of the error committed
in the computations.
These results allow us calibrate the adaptive discretization algorithm
separately for the gravitational potential, gravitational acceleration,
as well as the gravity gradient tensor components.

@@

We have implemented the calculation of
the tesseroid gravitational fields
with adaptive discretization
in version 1.2 of the open-source package \textit{Tesseroids}.
It is freely available online\footnote{
\url{http://tesseroids.leouieda.com}}\footnote{
\url{}\todo[inline]{Zenodo DOI}}
under the BSD 3-clause open-source license.
An archived version of the source code
is also available as part of this article.

\textit{Tesseroids} consists of command-line programs
written in the C programming language.
The package includes programs to calculate
the gravitational fields of tesseroids and
rectangular prisms (in both Cartesian and spherical coordinates),
as well as model and computation grid generation.
All programs receive input through
command-line arguments and the standard input channel (STDIN)
and output the results through the standard output channel (STDOUT).
For example,
the command to generate a regular grid with $NLON \times NLAT$ points,
calculate on it $g_z$ and $g_{zz}$ caused by
the tesseroids in a file ``MODELFILE'',
and save the results to a file called ``OUTPUT''
is:

@@

All figures and numerical experiments in this article
were produced in IPython notebooks\footnote{
\url{http://ipython.org}}
\citep{Perez2007}.
These files combine the analysis and figure generation source code
with text, equations, and the figures generated by the code.
We used the libraries
\textit{pandas} by \citet{Mckinney2010}
to perform the data analysis,
\textit{matplotlib} by \citet{Hunter2007} for graphs,
and \textit{Mayavi} by \citet{Ramachandran2011} for 3D graphs.

@@

We have implemented the values of the distance-size ratio producing $0.1\%$
accuracy determined above as defaults for the software \textit{Tesseroids}.
We chose the conservative value of $D=8$ for the gravity gradient components
as a fail-safe alternative.
Users can control the value of $D$ used in the computations through
command-line arguments to achieve greater performance at the cost of
accuracy.

@@

We have presented the open-source software \textit{Tesseroids}.
It consists of command-line programs,
written in the C programming language,
to perform the forward modeling of gravitational fields in spherical
coordinates.
The fields are calculated from a mass model composed of spherical prisms, the
so-called tesseroids.
The volume integrals of the gravitational fields of a tesseroid are solved
numerically using the Gauss-Legendre Quadrature (GLQ).
The GLQ approximates the volume integrals by weighted sums of point mass
effects.
The error of the GLQ integration increases as the computation point gets closer
to the tesseroid.
To counter this effect,
the accuracy of the GLQ integration can be increased by using more point
masses or by dividing each tesseroid into smaller ones.

@@

The values of the distance-size ratio determined above were
incorporated as defaults in the software \textit{Tesseroids}.
We chose the value $D=8$ for the gravity gradients as a conservative default.
If the user desires, the value of $D$ used can be controlled by a command-line
argument.

>>>>

\pdfoutput=1
\documentclass[paper,twocolumn,twoside]{geophysics}
%\documentclass[manuscript]{geophysics}
%\documentclass[manuscript,endfloat]{geophysics}
%\documentclass[twocolumn]{article}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\newcommand{\mitbf}[1]{\mathbf{{#1}}}

@@

We have implemented a modified version of the adaptive discretion of
\citet{Li2011} into the open-source software package \emph{Tesseroids}.
The software uses the Cartesian formula of \citet{Grombein2013} for improved
performance and robustness.
Previous versions of the software have been used by, e.g.,
\citet{Alvarez2012, Bouman2013, Bouman2013a, Mariani2013, Braitenberg2014,
Braitenberg2011, Fullea2014}.
This article describes the software design and the implementation
of our modified adaptive discretization algorithm.
We also present a numerical investigation of the error committed
in the computations.
These results allow us calibrate the adaptive discretization algorithm
separately for the gravitational potential, gravitational acceleration,
as well as the gravity gradient tensor components.

@@

We have implemented the calculation of
the tesseroid gravitational fields
with adaptive discretization
in version 1.2 of the open-source package \emph{Tesseroids}.
It is freely available online\footnote{
\url{http://tesseroids.leouieda.com}}\footnote{
\url{}\todo[inline]{Zenodo DOI}}
under the BSD 3-clause open-source license.
An archived version of the source code
is also available as part of this article.

\emph{Tesseroids} consists of command-line programs
written in the C programming language.
The package includes programs to calculate
the gravitational fields of tesseroids and
rectangular prisms (in both Cartesian and spherical coordinates),
as well as model and computation grid generation.
All programs receive input through
command-line arguments and the standard input channel (STDIN)
and output the results through the standard output channel (STDOUT).
For example,
the command to generate a regular grid with $NLON \times NLAT$ points,
calculate on it $g_z$ and $g_{zz}$ caused by
the tesseroids in a file ``MODELFILE'',
and save the results to a file called ``OUTPUT''
is:

@@

All figures and numerical experiments in this article
were produced in IPython notebooks\footnote{
\url{http://ipython.org}}
\citep{Perez2007}.
These files combine the analysis and figure generation source code
with text, equations, and the figures generated by the code.
We used the libraries
\emph{pandas} by \citet{Mckinney2010}
to perform the data analysis,
\emph{matplotlib} by \citet{Hunter2007} for graphs,
and \emph{Mayavi} by \citet{Ramachandran2011} for 3D graphs.

@@

We have implemented the values of the distance-size ratio producing $0.1\%$
accuracy determined above as defaults for the software \emph{Tesseroids}.
We chose the conservative value of $D=8$ for the gravity gradient components
as a fail-safe alternative.
Users can control the value of $D$ used in the computations through
command-line arguments to achieve greater performance at the cost of
accuracy.

@@

We have presented the open-source software \emph{Tesseroids}.
It consists of command-line programs,
written in the C programming language,
to perform the forward modeling of gravitational fields in spherical
coordinates.
The fields are calculated from a mass model composed of spherical prisms, the
so-called tesseroids.
The volume integrals of the gravitational fields of a tesseroid are solved
numerically using the Gauss-Legendre Quadrature (GLQ).
The GLQ approximates the volume integrals by weighted sums of point mass
effects.
The error of the GLQ integration increases as the computation point gets closer
to the tesseroid.
To counter this effect,
the accuracy of the GLQ integration can be increased by using more point
masses or by dividing each tesseroid into smaller ones.

@@

The values of the distance-size ratio determined above were
incorporated as defaults in the software \emph{Tesseroids}.
We chose the value $D=8$ for the gravity gradients as a conservative default.
If the user desires, the value of $D$ used can be controlled by a command-line
argument.
